Sparse Function-space Representation \\ of Neural Networks

### 2023-05-24 paper/main.tex
fixed small things

\section{\our for continual learning}
\label{app:cl}
This section provides additional details about the derivation of the \our-based regularizer described in \cref{sec:sequential} and how we extended it for dealing with the multi-output setting in our experiments.
\subsection{Derivation of the regularizer term}
For continual learning, the model cannot access the training data from previous tasks up to task $t-1$, denoted here as $\dataset_\textrm{old} = \bigcup_{s=1}^{t-1} \dataset_s$, and for each new task $t$ receives a new chunk of data $\dataset_\textrm{new} = \dataset_t$. On the other hand, in the supervised learning setting (described in \cref{sec:methods}), the neural network model has access to all the training data at once, \ie, the objective we are trying to optimize is
\begin{equation}
	\weights^* = \arg \min_\weights \mathcal{L}(\dataset_\textrm{old} \cup \dataset_\textrm{new}, \weights) = \arg \min_\weights \sum_{i \in \dataset_\textrm{old} \cup \dataset_\textrm{new}}\ell(f_\weights(\mathbf{x}_{i}), y_i) + \delta \mathcal{R}(\weights).
\end{equation} 
Our goal in the continual setting is to replace the missing data coming from $\dataset_\textrm{old}$ by resorting to a functional regularizer that accounts for that by forcing the predictions of the neural networks over the inducing points $\mathcal{Z}_t$ to come from $f^{\textrm{old}}_\weights(\vz_i) \sim q^\textrm{old}(\vu),  \forall \vz_i \in \mathcal{Z}_t$. As shown in \citep{khan2017conjugate, adam2021dual}, we can view the dual parameters derived in \cref{sec:sparse-dual-gp} as linked to approximate likelihood terms. Using this insight, we can rewrite the sparse GP posterior for $q(\vu)$ in terms of the following approximate normal distributions:
%
\begin{align} \label{eq:normal}
	q(\vu)
	& \propto  \Norm(0, \MKzz) \prod_{i \in \dataset_\textrm{old}} \exp \! \left(-\frac{\hat{\beta}_i}{2}(\tilde{y}_i - \va_i^\top \vu)^2 \right) \\
        & \propto p(\vu) \Norm(\tilde{\vy}; \vu, \MB_\vu),
\end{align}
% The first normals are the approximate likelihoods 
where $\va_i^\top = \vkzi^\top \MKzz^{-1}$ and $\tilde{y}_i = \hat{\alpha}_i + f_{\vw^*}(\vx_i)\hat{\beta}_i$.
The sparse GP posterior $q(\vu)$ is proportional to the product between the prior over the inducing points $p(\vu)$ and a Gaussian term $\Norm(\tilde{\vy}; \vu, \MB_\vu)$, which arises from the sparse approximation to the likelihood. Replacing $\va_i^\top \vu$ with $\vx_i^{\top}\vw$, we can see the similarity between the sparse GP and a linear model where the kernel is the linear kernel. We therefore use the key quadratic term from  $\Norm(\tilde{\vy}; \vu, \MB_\vu)$ as our regularizer as the missing data term, which after we take the $\log$ becomes $(\tilde{\vy} - \vu)^\top \bar{\MB}^{-1}(\tilde{\vy} - \vu)$.  Given that $\tilde{\vy}$ are in the function space we can instead us simply the previous neural network predictions $f^{\textrm{old}}_\weights(\vz_i)$ and save computations.

@@

\section{Experiment details}
\label{app:experiments}
%
We provide further details for the experiments presented in the main paper. The illustrative examples in \cref{fig:teaser,fig:banana} include toy problems for regression and classification. The 1D regression problem (\cref{fig:teaser}) uses an MLP with two hidden layers (64 hidden units each, with tanh activation), whereas the 2D banana binary classification problem uses an MLP with two hidden layers (64 hidden units each, with sigmoid activation). Both examples use Adam as the optimizer. We include Python notebooks for trying out the models. The illustrative HMC baseline in \cref{fig:banana} (right) was implemented with the help of hamiltorch (\url{https://github.com/AdamCobb/hamiltorch}). We used ten chains with 10,000 samples each and remove 2000 samples as burn-in.

>>>>

\section{\our for continual learning}
\label{app:cl}
This section provides additional details about the derivation of the \our-based regularizer described in \cref{sec:sequential} and how we extended it for dealing with the multi-output setting in our experiments.
\subsection{Derivation of the regularizer term}
For continual learning, the model cannot access the training data from previous tasks up to task $t-1$, denoted here as $\dataset_\textrm{old} = \bigcup_{s=1}^{t-1} \dataset_s$, and for each new task $t$ receives a new chunk of data $\dataset_\textrm{new} = \dataset_t$. On the other hand, in the supervised learning setting (described in \cref{sec:methods}), the neural network model has access to all the training data at once, \ie, the objective we are trying to optimize is
\begin{equation}
	\weights^* = \argmin_\weights \mathcal{L}(\dataset_\textrm{old} \cup \dataset_\textrm{new}, \weights) = \argmin_\weights \sum_{i \in \dataset_\textrm{old} \cup \dataset_\textrm{new}}\ell(f_\weights(\mathbf{x}_{i}), y_i) + \delta \mathcal{R}(\weights).
\end{equation} 
Our goal in the continual setting is to replace the missing data coming from $\dataset_\textrm{old}$ by resorting to a functional regularizer that accounts for that by forcing the predictions of the neural networks over the inducing points $\mathcal{Z}_t$ to come from $f^{\textrm{old}}_\weights(\vz_i) \sim q^\textrm{old}(\vu),  \forall \vz_i \in \mathcal{Z}_t$. As shown in \citep{khan2017conjugate, adam2021dual}, we can view the dual parameters derived in \cref{sec:sparse-dual-gp} as linked to approximate likelihood terms. Using this insight, we can rewrite the sparse GP posterior for $q(\vu)$ in terms of the following approximate normal distributions:
%
\begin{align} \label{eq:normal}
	q(\vu)
	& \propto  \Norm(\mathbf{0}, \MKzz) \prod_{i \in \dataset_\textrm{old}} \exp \! \left(-\frac{\hat{\beta}_i}{2}(\tilde{y}_i - \va_i^\top \vu)^2 \right) \\
        & \propto p(\vu) \Norm(\tilde{\vy}; \vu, \MB_\vu),
\end{align}
% The first normals are the approximate likelihoods 
where $\va_i^\top = \vkzi^\top \MKzz^{-1}$ and $\tilde{y}_i = \hat{\alpha}_i + f_{\vw^*}(\vx_i)\hat{\beta}_i$.
The sparse GP posterior $q(\vu)$ is proportional to the product between the prior over the inducing points $p(\vu)$ and a Gaussian term $\Norm(\tilde{\vy}; \vu, \MB_\vu)$, which arises from the sparse approximation to the likelihood. Replacing $\va_i^\top \vu$ with $\vx_i^{\top}\vw$, we can see the similarity between the sparse GP and a linear model, for which the kernel is the linear kernel. Therefore, we use the quadratic term from  $\Norm(\tilde{\vy}; \vu, \MB_\vu)$ as our regularizer in place of the missing data term. By taking the $\log(\cdot)$, the regularizer becomes $(\tilde{\vy} - \vu)^\top \bar{\MB}^{-1}(\tilde{\vy} - \vu)$.  Given that $\tilde{\vy}$ are in the function space we can instead us simply the previous neural network predictions $f^{\textrm{old}}_\weights(\vz_i)$ and save computations.

@@

\section{Experiment details}
\label{app:experiments}
%
We provide further details for the experiments presented in the main paper. In tables throughout the paper, we mark the best-performing method/setting by typesetting them boldface. Following this, we check with a two-tailed $t$-test whether the bolded entry significantly differs from the rest. Those not differing from the best performing one with a significance level of 5\% are also bolded. The illustrative examples in \cref{fig:teaser,fig:banana} include toy problems for regression and classification. The 1D regression problem (\cref{fig:teaser}) uses an MLP with two hidden layers (64 hidden units each, with tanh activation), whereas the 2D banana binary classification problem uses an MLP with two hidden layers (64 hidden units each, with sigmoid activation). Both examples use Adam as the optimizer. We include Python notebooks for trying out the models. The illustrative HMC baseline in \cref{fig:banana} (right) was implemented with the help of hamiltorch (\url{https://github.com/AdamCobb/hamiltorch}). We used ten chains with 10,000 samples each and remove 2000 samples as burn-in.
